Spatiotemporal transmission risk

### 2023-03-14 analysis/spatiotemporal-transmission-risk.Rmd
update spatiotemporal modeling workflow

# plot with and height of input and data graphs
aw <- 12
ah <- 8
```

@@

@@

```{r}
clinical_dta <- readRDS("../data-clean/Masi/clinical-data/clinical-data.rds")
share_tb <- clinical_dta %>%
  group_by(date, patient_id) %>%
  summarize(tb = any(tb_suspect == "yes")) %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(n = n(),
            n_tb = sum(tb)) %>%
  arrange(date) %>%
  mutate(prop_tb = n_tb / n * 100,
         date = format(date, format = "%b %d"),
         date = factor(date, levels = unique(date))) %>%
  ungroup() 
share_tb_pl <- share_tb %>%
  ggplot(aes(x = date, y = n_tb)) +
  geom_bar(stat = "identity", fill = "grey", color = "black", width = .5) +
  annotate("text", x = "Nov 09", y = mean(share_tb$n_tb), label = paste0("Mean = ",round(mean(share_tb$n_tb), 1)), color = "darkred", size = 8 / cm(1), vjust = -1) +
  geom_hline(aes(yintercept = mean(share_tb$n_tb)), color = "darkred") +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  scale_x_discrete(breaks = c("Oct 11", "Oct 18", "Oct 25", "Nov 02", "Nov 09", "Nov 16")) +
  labs(y = "No. of TB patients") +
  theme_bw2() +
  theme(axis.title.x = element_blank(),
        plot.margin = margin(5.5, 10, 5.5, 5.5))
save_plot(share_tb_pl, pdf_file = "../results/data/share-tb-patients.pdf", w = aw, h = ah)
```

@@

```{r}
co2_pl <- readRDS("../data-clean/Masi/environmental-data/co2-temp-humid.rds") %>%
  filter(location %in% c("waiting room", "tb room")) %>%
  mutate(location = factor(ifelse(location == "tb room", "TB room", "Waiting room"), levels = c("Waiting room", "TB room")),
         time = as.POSIXct(time, format = "%H:%M", tz = "CET")) %>%
  ggplot(aes(x = time, y = co2, color = location, fill = location)) +
  stat_lineribbon(alpha = .25, size = .5) +
  scale_x_datetime(expand = c(0,0), date_breaks = "2 hours", date_labels = "%H:%M") +
  scale_y_continuous(expand = c(0,0), breaks = scales::breaks_width(width = 100)) +
  scale_fill_brewer(palette = "Accent") +
  scale_color_brewer(palette = "Accent") +
  labs(y = expression("CO"[2]*" (ppm)"), fill = "Quantile", color = "Quantile") +
  theme_bw2() +
  theme(legend.position = c(.8,.95), legend.direction = "horizontal", legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.2, "cm"),
        plot.margin = margin(5.5, 10, 5.5, 5.5), axis.title.x = element_blank(), legend.title = element_blank()) 
save_plot(co2_pl, pdf_file = "../results/data/co2-levels.pdf", w = aw, h = ah)
```

@@

```{r}
# background image
clinic_img <- png::readPNG("../data-raw/Masi/building/clinic_clipped.png")
```

<!-- ```{r} -->
<!-- # choose cell size such that the squared cell has 1m3 volume -->
<!-- cellVolume <- 1 -->
<!-- cellSize <- sqrt(cellVolume * 1e9 / 3e3)  -->

<!-- # waiting room -->
<!-- sh_wr <- vect("../data-raw/Masi/building/waiting_room.shp") -->
<!-- sf_wr <- st_as_sf(sh_wr) -->
<!-- wrRoomSP <- shapeToSpatial(sh_wr, cellSize) -->
<!-- wrVol <- compute_volume(wrRoomSP) -->
<!-- wrCellCoords <- create_coord_df(wrRoomSP) -->
<!-- wrRoomXY <- as.matrix(dplyr::select(wrCellCoords, x, y)) -->

<!-- # tb room -->
<!-- sh_tb <- vect("../data-raw/Masi/building/tb_room.shp") -->
<!-- sf_tb <- st_as_sf(sh_tb) -->
<!-- tbRoomSP <- shapeToSpatial(sh_tb, cellSize) -->
<!-- tbVol <- compute_volume(tbRoomSP) -->
<!-- tbCellCoords <- create_coord_df(tbRoomSP) -->
<!-- tbRoomXY <- as.matrix(dplyr::select(tbCellCoords, x, y)) -->

<!-- # passage -->
<!-- sh_pa <- vect("../data-raw/Masi/building/passage.shp") -->
<!-- sf_pa <- st_as_sf(sh_pa) -->
<!-- paRoomSP <- shapeToSpatial(sh_pa, cellSize) -->
<!-- paVol <- compute_volume(paRoomSP) -->
<!-- paCellCoords <- create_coord_df(paRoomSP) -->
<!-- paRoomXY <- as.matrix(dplyr::select(paCellCoords, x, y)) -->

<!-- # combine -->
<!-- clinicCellCoords <- rbind( -->
<!--   mutate(wrCellCoords, location = "waiting room"), -->
<!--   mutate(tbCellCoords, location = "tb room"), -->
<!--   mutate(paCellCoords, location = "passage") -->
<!-- ) %>% mutate(id = 1:nrow(.)) -->
<!-- clinicCellCoordsXY <- cbind(clinicCellCoords$x, clinicCellCoords$y) -->
<!-- ``` -->

@@

```{r}
files <- list.files("../data-clean/Masi/patient-tracking-data/", full.names = T)
no_people <- do.call(rbind, lapply(files, function(x) readRDS(paste0(x, "/patient-id-data.rds")))) %>%
  mutate(date = format(time, format = "%Y-%m-%d"),
         time = as.POSIXct(time, format = "%H:%M", tz = "CET"),
         daytime = ifelse(time <= "2021-10-25 12:00:00", "Morning", "Afternoon"),
         daytime = factor(daytime, levels = c("Morning", "Afternoon")))

# over time
no_people_time_pl <- no_people %>%
  group_by(date, time) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  ggplot(aes(x = time, y = n)) +
  geom_line() +
  #stat_lineribbon(alpha = .25, size = .5) +
  scale_x_datetime(expand = c(0,0), date_breaks = "2 hours", date_labels = "%H:%M") +
  scale_y_continuous(expand = c(0,0), limits = c(0, NA), breaks = scales::breaks_width(5)) +
  scale_fill_brewer(palette = "OrRd") +
  labs(y = "No. of clinical attendees") +
  theme_bw2() +
  theme(axis.title.x = element_blank())
  #labs(y = "No. of clinical attendees", fill = "Quantile", color = "Quantile") +
  # theme(legend.position = c(.8,.9), legend.direction = "horizontal", legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.2, "cm"),
  #       plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  # guides(fill = guide_legend(title.position = "top"))
save_plot(no_people_time_pl, pdf_file = "../results/data/no-people-over-time.pdf", w = aw, h = ah)

# Spatial
no_people_spatial_pl <- no_people %>%
  dplyr::select(x, y) %>%
  adespatial::rotation(angle = -8*pi/180) %>%
  as.data.frame() %>%
  set_names(c("x", "y")) %>%
  mutate(x = -x) %>%
  ggplot(aes(x = x, y = y)) +
  geom_bin2d(binwidth = 100) +
  ggpubr::background_image(clinic_img) +
  annotate("text", x = -13200, y = -5300, label = "Entrance", size = 8 / cm(1), fontface = 2, hjust=0,vjust=0) +
  annotate("text", x = -13400, y = -2300, label = "Waiting room", size = 8 / cm(1), fontface = 2, hjust=0,vjust=0) +
  annotate("text", x = -5000, y = -4100, label = "Registration", size = 8 / cm(1), fontface = 2, angle = 90, hjust=0,vjust=0) +
  annotate("text", x = 2600, y = -2700, label = "TB room", size = 8 / cm(1), fontface = 2) +
  scale_fill_stepsn(colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd")[c(1,3,5,7,9)], breaks = c(1,5,10,50,100), limits = c(0,100)) +
  scale_y_continuous(expand = expansion(mult = c(0,.18))) +
  scale_x_continuous(expand = expansion(mult = c(0,0.025))) +
  labs(fill = "Density") +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = c(0.825,0.075), legend.direction = "horizontal", legend.title = element_blank(), #legend.title = element_text(vjust = .75),
        panel.background = element_rect(fill = RColorBrewer::brewer.pal(n = 3, name = "YlOrRd")[1]), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
save_plot(no_people_spatial_pl, pdf_file = "../results/data/no-people-spatial.pdf", w = aw, h = ah)
```

<!-- ```{r} -->
<!-- # number of people in space -->
<!-- nDF <- tibble() -->
<!-- for (tp in time_period) { -->
<!--   n_tp <- clinXov %>% -->
<!--     filter(between(time, tp, tp + seconds(59))) %>% -->
<!--     mutate(location = mapply(function(x,y) clinicCellCoords$location[find_raster(x,y,clinicCellCoordsXY)], x, y)) %>% -->
<!--     group_by(person_id, location) %>% -->
<!--     slice(1) %>% -->
<!--     ungroup() %>% -->
<!--     group_by(location) %>% -->
<!--     summarize(n = n()) %>% -->
<!--     ungroup()  -->
<!--   for (loc in unique(clinicCellCoords$location)) { -->
<!--     if (!(loc %in% n_tp$location)) { -->
<!--       n_tp <- rbind(n_tp, data.frame(location = loc, n = 0)) -->
<!--     }  -->
<!--   } -->
<!--   nDF <- rbind(nDF, n_tp %>% mutate(time = tp)) -->
<!-- } -->
<!-- nDF <- nDF %>%  -->
<!--   mutate(time = as.POSIXct(time, origin = as.Date("1970-01-01"))) -->
<!-- nDF$n[nDF$location=="waiting room"] <- nDF$n[nDF$location=="waiting room"] + nDF$n[nDF$location=="passage"] -->
<!-- nDF$n[nDF$location=="passage"] <- nDF$n[nDF$location=="waiting room"] + nDF$n[nDF$location=="passage"] -->
<!-- ggplot(nDF, aes(x = time, y = n, color = location)) + geom_line() -->
<!-- ``` -->

<!-- ### Movement -->

<!-- ```{r} -->
<!-- patMov <- clinXov %>% -->
<!--   mutate(id = map2_int(x, y, find_raster, cellCoordsXY = clinicCellCoordsXY)) -->
<!-- ``` -->

@@

## Simulation

```{r}
# Settings
set.seed(12345)
nSim <- 100

# Input parameters
inputPar <- data.frame(
  spatial_mu = rMuIc(nSim),
  quanta = rq(nSim),
  V = rV(nSim),
  co2_ex = co2_exhaled,
  co2_out = min(co2$co2),
  co2_gen = rG(nSim),
  unmasked_TB = rTBunmasked(nSim, nTBmasked)
)

# Loop
for (s in 1:nSim) {
  message(sprintf("Simulation %i of %i", s, nSim))

  # parameters
  smu <- inputPar$spatial_mu[s]
  q <- inputPar$quanta[s]
  V <- inputPar$V[s]
  Ca <- inputPar$co2_ex[s]
  Co <- inputPar$co2_out[s]
  G <- inputPar$co2_gen[s]
  unmasked_TB <- inputPar$unmasked_TB[s]

  # transformed parameters
  # quanta emission rate per minute
  ER <- q / 60

  # TB patients
  masked_tb <- yesTB$person_id
  unmasked_tb <- sample(noTB$person_id, size = unmasked_TB)
  TBpatients <- patMov %>%
    filter(person_id %in% c(masked_tb, unmasked_tb)) %>%
    dplyr::select(person_id, observation_id, time, x, y) 
  TBpatients_full <- TBpatients %>%
    group_by(person_id) %>%
    summarize(time = list(seq(min(time), max(time), by = "sec"))) %>%
    ungroup() %>%
    unnest(cols = c(time))
  TBpatients_full <- TBpatients_full %>%
    left_join(TBpatients, by = c("person_id", "time")) %>%
    tidyr::fill(observation_id, x, y, .direction = "down")

  # no TB patients
  noTBpatients <- patMov %>%
    filter(!(person_id %in% c(masked_tb, unmasked_tb))) %>%
    dplyr::select(person_id, time, id) 
  noTBpatients_full <- noTBpatients %>%
    group_by(person_id) %>%
    summarize(time = list(seq(min(time), max(time), by = "sec"))) %>%
    ungroup() %>%
    unnest(cols = c(time))
  noTBpatients_full <- noTBpatients_full %>%
    left_join(noTBpatients, by = c("person_id", "time")) %>%
    tidyr::fill(id, .direction = "down") %>%
    mutate(time = as.POSIXct(format(time, "%Y-%m-%d %H:%M"))) %>%
    group_by(person_id, id, time) %>%
    summarize(minutes = n() / 60) %>%
    ungroup()

  # df to store quanta concentration by minute
  quantaDF <- clinicCellCoords %>%
    mutate(time = time_period[1] - minutes(1),
           I_SP = 0,
           I_WM = 0,
           E_SP = 0,
           E_WM = 0,
           N_SP = 0,
           AER = NA,
           RR = NA,
           N_WM = 0,
           N_SP = 0)

  for (t in 1:length(time_period)) {

    # time point
    tp <- time_period[t] 
    quantaDF_new <- clinicCellCoords %>%
      mutate(time = tp,
             I_SP = 0,
             I_WM = 0,
             E_SP = 0,
             E_WM = 0,
             N_SP = 0,
             N_WM = 0)

    # spatial spread of infectious individuals and quanta emission
    TBpatients_tp <- TBpatients_full %>%
      dplyr::filter(between(time, tp, tp + minutes(1)))
    if (nrow(TBpatients_tp) > 0) {
      TBpatients_tp <- TBpatients_tp %>%
        nest(data = -person_id)
      # iterate over patients
      for (p in 1:nrow(TBpatients_tp)) {
        # duration and location
        duration <- nrow(TBpatients_tp$data[[p]])
        xy <- TBpatients_tp$data[[p]] %>% dplyr::select(x, y) %>% slice(1) %>% unlist()
        # generated quanta by cell
        locationID <- find_raster(xy[1], xy[2], clinicCellCoordsXY)
        locationName <- clinicCellCoords$location[locationID]
        locationCellCoords <- filter(clinicCellCoords, location == locationName)
        locationShape <- if(locationName == "waiting room") {sf_wr} else if(locationName == "tb room") {sf_tb} else {sf_pa}
        cellInReach <- numeric(nrow(quantaDF_new))
        cellDistance <- numeric(nrow(quantaDF_new))
        isIinCell <- ifelse(quantaDF_new$location==locationName, 1, 0 )
        for (cell in 1:nrow(quantaDF_new)) {
          # if cell is in room/location
          if (quantaDF_new$location[cell] != locationName) {
            cellInReach[cell] <- 0
            cellDistance[cell] <- -999
          } else {
            # can aerosols travel to this cell?
            straight_line <- st_linestring(rbind(xy, c(quantaDF_new$x[cell], quantaDF_new$y[cell]))) 
            intersects <- st_difference(straight_line, locationShape)
            cellInReach[cell] <- ifelse(length(intersects) == 0, 1, 0)
            cellDistance[cell] <- convert_dist(st_length(straight_line))
          }
        }
        # proportion of infectious individuals per cell per minute
        distance_dens <- cellInReach * dIc(cellDistance, smu)
        quantaDF_new$I_SP <- quantaDF_new$I_SP + distance_dens / sum(distance_dens) * (duration / 60)
        quantaDF_new$I_WM <- quantaDF_new$I_WM + isIinCell / sum(isIinCell)
      }
      # quanta emission
      quantaDF_new$E_SP <- quantaDF_new$E_SP + ER * quantaDF_new$I_SP
      quantaDF_new$E_WM <- quantaDF_new$E_WM + ER * quantaDF_new$I_WM
    }

    # compute air exchange rate and quanta removal
    Ct <- co2 %>%
      group_by(location) %>%
      # if CO2 is missing, take the last observed value
      filter(date_time <= tp) %>%
      arrange(desc(date_time)) %>%
      slice(1) %>%
      mutate(date_time = tp) %>%
      ungroup() %>%
      # add number of people
      left_join(filter(nDF, time == tp), by = c("date_time" = "time", "location")) %>%
      # add volume
      mutate(vol = ifelse(location == "tb room", tbVol, wrVol + paVol)) %>%
      # compute AER and RR per minute
      mutate(AER = compute_AER(G, Co, co2, n, vol),
             RR = compute_RR(AER) / 60) 
    quantaDF_new <- quantaDF_new %>% 
      left_join(Ct %>% dplyr::select(location, AER, RR), by = "location")

    # previous quanta concentration
    N0_SP <- quantaDF %>% 
      filter(time == tp - minutes(1)) %>%
      dplyr::select(N_SP) %>% unlist
    N0_WM <- quantaDF %>% 
      filter(time == tp - minutes(1)) %>%
      dplyr::select(N_WM) %>% unlist

    # compute quanta concentration
    quantaDF_new$N_SP <- pmap_dbl(list(E = quantaDF_new$E_SP, 
                                       RR = quantaDF_new$RR, 
                                       N0 = N0_SP,
                                       V = cellVolume), compute_Nt)
    quantaDF_new$N_WM <- pmap_dbl(list(E = quantaDF_new$E_WM, 
                                       RR = quantaDF_new$RR, 
                                       N0 = N0_WM,
                                       V = cellVolume), compute_Nt)

    # append new data
    quantaDF <- rbind(quantaDF, quantaDF_new)
  }

  # compute (spatio)temporal transmission risk for each individual
  risk_of_infection <- noTBpatients_full %>%
    left_join(quantaDF, by = c("id", "time")) %>%
    group_by(person_id) %>%
    summarize(exposure_time = sum(minutes),
              P_SP = 1-exp(-sum(N_SP * minutes)),
              P_WM = 1-exp(-sum(N_WM * minutes))) %>%
    ungroup()

  # compute average transmission risk
  sum_I <- data.frame(location = c("waiting room", "passage", "tb room"),
                      I = length(c(masked_tb, unmasked_tb))) 
  mean_n <- nDF %>% 
    group_by(location) %>%
    summarize(n = median(n) + 1) %>%
    ungroup()
  mean_f <- co2 %>%
    group_by(location) %>%
    summarize(f = (mean(co2) - Co) / Ca) %>%
    ungroup()
  risk_of_infection_av <- noTBpatients_full %>%
    left_join(quantaDF %>% dplyr::select(id, time, location), by = c("id", "time")) %>%
    group_by(person_id, location) %>%
    summarize(minutes = sum(minutes)) %>%
    ungroup() %>%
    left_join(sum_I, by = "location") %>%
    left_join(mean_n, by = "location") %>%
    left_join(mean_f, by = "location") %>%
    mutate(mu = f * I * ER * minutes / n) %>%
    group_by(person_id) %>%
    summarize(P_SS = 1 - exp(-sum(mu))) %>%
    ungroup()
  risk_of_infection <- risk_of_infection %>%
    left_join(risk_of_infection_av, by = "person_id")

  # save simulation
  saveRDS(quantaDF, paste0("../simulations/2021-10-25/sim-quanta", s, ".rds"))
  saveRDS(risk_of_infection, paste0("../simulations/2021-10-25/sim-transRisk", s, ".rds"))
}
```

@@

### TB risk

```{r}
# files
transRiskFiles <- list.files("../simulations/2021-10-25", full.names = T)
transRiskFiles <- transRiskFiles[grepl("transRisk", transRiskFiles)]

# read files
transRiskDF <- data.frame(file = transRiskFiles) %>%
  mutate(simulation = stringi::stri_extract(basename(file), regex = "\\d{1,4}"), 
         data = lapply(file, readRDS)) %>%
  unnest(cols = c(data))

# QQ Plot
transRiskDF_Q <- transRiskDF %>%
  dplyr::select(simulation, P_SP, P_WM) %>%
  melt("simulation") %>%
  group_by(simulation, variable) %>%
  summarize(quantile = seq(0.05, .95, 0.05),
            value = quantile(value, seq(0.05, 0.95, 0.05))) %>%
  ungroup()

transRiskDF_Q_av <- transRiskDF_Q %>%
  group_by(variable, quantile) %>%
  summarize(value = mean(value)) %>%
  ungroup()

transRiskDF_Q_av %>%
  ggplot(aes(x = quantile, y = value, color = variable)) +
  geom_line()

# average transmission risk across patients
average_tr <- transRiskDF %>%
  dplyr::select(simulation, P_SP, P_WM) %>%
  mutate(P_SP = P_SP * 100,
         P_WM = P_WM * 100) %>%
  melt("simulation") %>%
  group_by(simulation, variable) %>%
  summarize(`Mean` = mean(value),
            `Standard deviation` = sd(value),
            `Median` = median(value),
            `Tail risk (95%-Quantile)` = quantile(value, .95)) %>%
  ungroup()

figA <- average_tr %>%
  rename(model = variable) %>%
  mutate(model = ifelse(model == "P_SP", "Spatiotemporal modeling", "Temporal modeling")) %>%
  melt(c("simulation", "model")) %>%
  ggplot(aes(x = value, y = model, fill = model, color = model)) +
  stat_dots(quantiles = 100) +
  facet_wrap(~ variable, scales = "free_x") +
  scale_color_manual(values = cbPalette[2:3]) +
  scale_fill_manual(values = cbPalette[2:3]) +
  labs(x = "Risk of infection (%)", 
       title = "Fig. 1a: Risk of infection across patients",
       subtitle = "Colored dots show individual simulation results.") +
  theme_bw2() +
  theme(legend.position = "top", axis.title.y = element_blank(), axis.text.y = element_blank(), legend.title = element_blank(),
        legend.justification='left', legend.direction='horizontal')

figA

average_tr %>%
  dplyr::select(-simulation) %>%
  group_by(variable) %>%
  median_qi() %>%
  mutate_if(is.numeric, round, 2)

>>>>

# plot with and height of input and data graphs
width_descr <- 10.5 / cm(1)
height_descr <- 7.5 / cm(1)
text_descr <- 12
```

@@

```{r}
selected_dates <- as.Date(list.files("../data-clean/Masi/patient-tracking-data")[-1])
```

@@

```{r}
clinical_dta <- readRDS("../data-clean/Masi/clinical-data/clinical-data.rds") %>%
  filter(date %in% selected_dates)

# number of registered patients
clinical_dta %>%
  group_by(patient_id) %>%
  slice(1) %>%
  nrow()

# share of TB patients
share_tb <- clinical_dta  %>%
  group_by(date, patient_id) %>%
  summarize(tb = any(tb_suspect == "yes")) %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(n = n(),
            n_tb = sum(tb)) %>%
  arrange(date) %>%
  mutate(prop_tb = n_tb / n * 100,
         date = format(date, format = "%b %d"),
         date = factor(date, levels = unique(date))) %>%
  ungroup() 
share_tb_pl <- share_tb %>%
  ggplot(aes(x = date, y = n_tb)) +
  geom_bar(stat = "identity", fill = "grey", color = "black", width = .5) +
  annotate("text", x = "Nov 10", y = mean(share_tb$n_tb), label = paste0("Mean = ",round(mean(share_tb$n_tb), 1)), color = "darkred", size = 8 / cm(1), vjust = -1) +
  geom_hline(aes(yintercept = mean(share_tb$n_tb)), color = "darkred") +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) +
  scale_x_discrete(breaks = format(selected_dates, format = "%b %d")[c(1, 4, 8, 12)]) +
  labs(y = "No. of masked TB patients") +
  theme_classic() +
  theme(axis.title.x = element_blank(),
        plot.margin = margin(5.5, 5.5, 5.5, 5.5),
        text = element_text(size = text_descr))
ggsave(plot = share_tb_pl, filename = "../doc/IWHOD/presentation/figures/share-tb-patients.png", w = width_descr, h = height_descr)
```

@@

```{r}
co2_pl <- readRDS("../data-clean/Masi/environmental-data/co2-temp-humid.rds") %>%
  filter(location %in% c("waiting room", "tb room"),
         date %in% selected_dates) %>%
  mutate(location = factor(ifelse(location == "tb room", "TB room", "Waiting room"), levels = c("Waiting room", "TB room")),
         time = as.POSIXct(time, format = "%H:%M", tz = "CET")) %>%
  ggplot(aes(x = time, y = co2, color = location, fill = location)) +
  stat_lineribbon(alpha = .25, size = .75) +
  scale_x_datetime(expand = c(0,0), date_breaks = "2 hours", date_labels = "%H:%M") +
  scale_y_continuous(expand = c(0,0), breaks = scales::breaks_width(width = 100)) +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  scale_color_manual(values = wes_palette("Moonrise2")) +
  labs(y = expression("CO"[2]*" (ppm)"), fill = "Quantile", color = "Quantile") +
  theme_classic() +
  theme(legend.position = c(.65,.95), legend.direction = "horizontal", legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.2, "cm"),
        plot.margin = margin(5.5, 5.5, 5.5, 5.5), axis.title.x = element_blank(), legend.title = element_blank(),
        text = element_text(size = text_descr)) 
ggsave(plot = co2_pl, filename = "../doc/IWHOD/presentation/figures/co2-levels.png", width = width_descr, height = height_descr)
```

@@

```{r}
# background image
clinic_img <- png::readPNG("../data-raw/Masi/building/clinic_clipped.png")

# building
cellSize <- sqrt(1 * 1e9 / 3e3) 
building <- terra::vect(paste("../data-raw", "Masi", "building", "clinic-vector.gpkg", sep = "/"))
building_sf <- sf::st_as_sf(building, crs = NA)
sf::st_crs(building_sf) <- NA
waiting_room <- shapeToSpatial(building_sf$geometry[2], cellSize)
waiting_room_df <- fortify(waiting_room)
passage <- shapeToSpatial(building_sf$geometry[4], cellSize)
passage_df <- fortify(passage)
tb_room <- shapeToSpatial(building_sf$geometry[3], cellSize)
tb_room_df <- fortify(tb_room)

# test
waiting_room_df$value <- rnorm(nrow(waiting_room_df))
passage_df$value <- rnorm(nrow(passage_df))
tb_room_df$value <- rnorm(nrow(tb_room_df))

ggplot(mapping = aes(x = long, y = lat, group = group, fill = value)) +
  geom_polygon(data = waiting_room_df) +
  geom_polygon(data = passage_df) +
  geom_polygon(data = tb_room_df) +
  ggpubr::background_image(clinic_img) +
  annotate("text", x = -13200, y = -5300, label = "Entrance", size = 8 / cm(1), fontface = 2, hjust=0,vjust=0) +
  annotate("text", x = -13400, y = -2300, label = "Waiting room", size = 8 / cm(1), fontface = 2, hjust=0,vjust=0) +
  annotate("text", x = -5000, y = -4100, label = "Registration", size = 8 / cm(1), fontface = 2, angle = 90, hjust=0,vjust=0) +
  annotate("text", x = 2600, y = -2700, label = "TB room", size = 8 / cm(1), fontface = 2) +
  scale_y_continuous(expand = expansion(add=c(2500, 4000))) +
  scale_x_continuous(expand = expansion(add=c(400, 500))) +
  labs(fill = "Density") +
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = c(0.825,0.075), legend.direction = "horizontal", legend.title = element_blank(), #legend.title = element_text(vjust = .75),
        panel.background = element_rect(fill = RColorBrewer::brewer.pal(n = 3, name = "YlOrRd")[1]), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

@@

```{r}
files <- list.files("../data-clean/Masi/patient-tracking-data/", full.names = T)[-1]
no_people <- do.call(rbind, lapply(files, function(x) readRDS(paste0(x, "/patient-id-data_matched-filtered.rds")))) %>%
  mutate(date = as.Date(format(time, format = "%Y-%m-%d")),
         time_min = as.POSIXct(format(time, format = "%H:%M"), format = "%H:%M"))

# number of clinical attendees
no_people %>%
  group_by(date, patient_id) %>%
  slice(1) %>%
  nrow()

# over time
no_people_time <- no_people %>%
  group_by(date, time_min, time) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(date, time_min) %>%
  summarize(n = mean(n)) %>%
  ungroup()

no_people_time_pl <- no_people_time %>%
  ggplot(aes(x = time_min, y = n)) +
  stat_lineribbon(alpha = .25, size = .5) +
  scale_x_datetime(expand = c(0,0), date_breaks = "2 hours", date_labels = "%H:%M") +
  scale_y_continuous(expand = c(0,0), limits = c(0, NA), breaks = scales::breaks_width(5)) +
  scale_fill_brewer(palette = "OrRd") +
  labs(y = "No. of clinical attendees", fill = "Quantile", color = "Quantile") +
  theme_bw2() +
  theme(axis.title.x = element_blank(), legend.position = c(.8,.9), legend.direction = "horizontal", legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.2, "cm"),
        plot.margin = margin(5.5, 10, 5.5, 5.5)) +
  guides(fill = guide_legend(title.position = "top"))
no_people_time_pl

# Spatial
no_people_spatial_pl <- no_people %>%
  ggplot(aes(x = x, y = y)) +
  geom_bin2d(binwidth = 100) +
  scale_fill_stepsn(colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"), trans = "log", breaks = c(1, 10, 100, 1000, 10000)) +
    ggpubr::background_image(clinic_img) +
    annotate("text", x = -13200, y = -5800, label = "Entrance", size = text_descr / cm(1), hjust=0,vjust=0) +
    annotate("text", x = -14000, y = -2600, label = "Waiting room", size = text_descr / cm(1), hjust=0,vjust=0) +
    annotate("text", x = -5800, y = -4500, label = "Registration", size = text_descr / cm(1), angle = 90, hjust=0,vjust=0) +
    annotate("text", x = 3000, y = -2700, label = "TB room", size = text_descr / cm(1)) +
    scale_y_continuous(expand = expansion(mult=c(0.05, 0.15))) +
    scale_x_continuous(expand = expansion(mult=c(0.03, 0))) +
    labs(fill = "Density") +
    theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(), legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.3, "cm"),
          legend.position = c(0.825,0.075), legend.direction = "horizontal",legend.title = element_text(vjust = .75),
          panel.background = element_rect(fill = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd")[1]), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(title.position = "top", label.position = "bottom"))
ggsave(plot=no_people_spatial_pl, filename = "../doc/IWHOD/presentation/figures/no-people-spatial.png", width = width_descr, height = height_descr)
```

@@

@@

```{r}
sim_results <- list.files("../simulations/Masi/2021-10-25/", full.names = T)
sim_results <- do.call(rbind, lapply(sim_results, readRDS))
```

### Transmission risk

```{r}
# individual patient transmission risk by simulation
transRisk <- sim_results %>%
  group_by(patient_id, sim) %>%
  slice(1) %>%
  ungroup()

# Average transmission risk across patients
transRisk_av <- transRisk %>%
  group_by(sim) %>%
  summarize(across(c(P_N, P_N_ns), mean)) %>%
  ungroup()

mean_risk <- round(mean(transRisk_av$P_N * 100), 1)
q25_risk <- round(quantile(transRisk_av$P_N * 100, .25), 1)
q75_risk <- round(quantile(transRisk_av$P_N * 100, .75), 1)
message(sprintf("Spatiotemporal transmission risk across clinical attendee: %s percent (IQR: %s to %s)", mean_risk, q25_risk, q75_risk))

# individual transmission risk across simulations
transRisk_indv <- transRisk %>%
  group_by(patient_id) %>%
  summarize(across(c(P_N, P_N_ns), list(mean = mean, median = median,q25 = function(x) quantile(x, .25), q75 = function(x) quantile(x, .75)))) %>%
  ungroup()

transRisk_indv_pl <- transRisk_indv %>%
  mutate(patient_id = factor(patient_id)) %>%
  ggplot(aes(x = reorder(patient_id, -P_N_mean), y = P_N_mean)) +
  geom_point(shape = 4, alpha = .5) +
  geom_hline(aes(yintercept = mean_risk / 100), color = "darkred") +
  geom_hline(aes(yintercept = q25_risk / 100), color = "darkred", linetype = "dashed") +
  geom_hline(aes(yintercept = q75_risk / 100), color = "darkred", linetype = "dashed") +
  annotate("text", x = 375, y = 0.0075, label = paste0("Average across patients = ", mean_risk, "% (IQR:", q25_risk, "-", q75_risk,")"),
           size = 8 / cm(1), color = "darkred") +
  scale_y_continuous(labels = function(x) x * 100, expand = expansion(mult = c(0., .05)), breaks = c(seq(0, 0.05, 0.01), 0.004)) +
  labs(y = "Mean risk of infection (%)", x = "Clinic attendee") +
  theme_bw2() +
  theme(axis.text.x = element_blank())
save_plot(transRisk_indv_pl, pdf_file = "../results/modeling/risk-of-infection.pdf", w = aw, h = ah)
